---
title: "National Intelligence Quotient (IQ) Scores"
author: "Peyton Hall"
date: "08/06/2024"
output:
  pdf_document:
    latex_engine: lualatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load Necessary Libraries
```{r Load Necessary Libraries}
library(readr)
library(ggplot2)
library(rworldmap)
library(RColorBrewer)
library(dplyr)
```


Read the CSV file
```{r Read the CSV file}
avgIQpercountry <- read_csv("~/Desktop/NationalIQScores/avgIQpercountry.csv")
avgIQpercountry
```


Let μ = Average National IQ, AF = Africa, AS = Asia, CA = Central America, 
EU = Europe, EA = Eurasia, NA = North America, OC = Oceania, SA = South America

Let α = 0.05 (significance level)

$H_0: \mu_\text{AF} = \mu_\text{AS} = \mu_\text{CA} = \mu_\text{EU} = \mu_\text{EA} = \mu_\text{NA} = \mu_\text{OC} = \mu_\text{SA}$ 
vs 
$H_a: \text{At least two means differ}$


Hypothesis Testing
```{r Hypothesis Testing}
# Perform a One-Way Analysis of Variance (ANOVA)
iq_aov <- aov(`Average IQ`~Continent, data = avgIQpercountry)
summary(iq_aov)
TukeyHSD(iq_aov) # pairwise comparison test
```
f = 33.93; p-value = <2e-16 (i.e. 0.0000000000000002)

Reject H0; there is no evidence to support the claim that the average IQ among
each continent is equal. Viz., there is a 0.00000000000002% probability this 
difference occurred by chance. The difference occurred through observations of 
IQ scores.

In the pairwise comparison test, it is found the following continents are 
significantly different in average IQ scores:
Asia-Africa, Europe-Africa, Europe/Asia-Africa, North America-Africa, 
Oceania-Africa, South America-Africa, Central America-Asia, Europe-Asia, 
Europe-Central America, North America-Central America, Oceania-Central America,
South America-Europe.

The following continents are not significantly different in average IQ scores:
Central America-Africa, Europe/Asia-Asia, North America-Asia, Oceania-Asia,
South America-Asia, Europe/Asia-Central America, South America-Central America,
Europe/Asia-Europe, North America-Europe, Oceania-Europe, 
North America-Europe/Asia, Oceania-Europe/Asia, South America-Europe/Asia, 
Oceania-North America, South America-North America, South America-Oceania.

Pairwise comparison determinations are based upon the "p adj" column. 
If p adj < 0.05, there is a significant difference. 
If p adj > 0.05, there is no significant difference.
That is not to say there are no smaller differences.


Define Global Variables for Graphing
```{r Define Global Variables for Graphing}
# Save the value of the column header "Average IQ" to the variable Average_IQ
Average_IQ <- avgIQpercountry$`Average IQ`

# Pass Average_IQ into the iq_colors function
iq_colors <- function(Average_IQ) {
  if (is.na(Average_IQ)) {
    return("gray")
    }
  else if (Average_IQ < 50) {
    return("#8B0000")
    }
  else if (50 <= Average_IQ & Average_IQ < 60) {
    return("#FF4500")
    }
  else if (60 <= Average_IQ & Average_IQ < 70) {
    return("#FFA500")
    }
  else if (70 <= Average_IQ & Average_IQ < 80) {
    return("#F5DEB3")
    }
  else if (80 <= Average_IQ & Average_IQ < 90) {
    return("#ADD8E6")
    }
  else if (90 <= Average_IQ & Average_IQ < 100) {
    return("#7B68EE")
  }
  else {
    return("#0000CD")
    }
} # end iq_colors

iq_labels <- function(Average_IQ) {
  if (is.na(Average_IQ)) {
    return("N/A")
    }
  else if (Average_IQ < 50) {
    return("<50") 
    }
  else if (50 <= Average_IQ & Average_IQ < 60) {
    return("50-59")
    }
  else if (60 <= Average_IQ & Average_IQ < 70) {
    return("60-69")
    }
  else if (70 <= Average_IQ & Average_IQ < 80) {
    return("70-79")
    }
  else if (80 <= Average_IQ & Average_IQ < 90) {
    return("80-89")
    }
  else if (90 <= Average_IQ & Average_IQ < 100) {
    return("90-99")
  }
  else {
    return("100+")
    }
} # end iq_labels

iq_breaks <- c(-Inf, 50, 60, 70, 80, 90, 100, Inf) # Define the breaks
```


Global Map of Average IQ per Country
```{r Global Map of Average IQ per Country}
# Set graphical parameters to use Times New Roman for the title
par(family = "serif")

# Join the country data to the map
sPDF <- joinCountryData2Map(avgIQpercountry, joinCode = "NAME", 
                            nameJoinColumn = "Country")

# Apply the color function to create a color palette for the map
sPDF$color <- sapply(sPDF$`Average IQ`, iq_colors)

# Define the legend text based on the IQ values
legend_labels <- unique(sapply(sort(unique(sPDF$`Average IQ`)), iq_labels))
legend_colors <- unique(sapply(sort(unique(sPDF$`Average IQ`)), iq_colors))

# Plot the map
mapParams <- mapCountryData(sPDF, 
                            nameColumnToPlot = "Average IQ", 
                            mapTitle = "Global Map of Average IQ per Country",
                            colourPalette = legend_colors, 
                            catMethod = iq_breaks,
                            oceanCol = "#FFFFFF", 
                            missingCountryCol = "gray",
                            addLegend = FALSE,
                            borderCol = "#000000",
                            xlim = c(-180, 180), ylim = c(-90, 90))

# Add "N/A" to the legend manually
legend_labels <- c(legend_labels, "N/A")
legend_colors <- c(legend_colors, "gray")

# Add the legend manually at the bottom
legend("bottom", 
       legend = legend_labels, 
       fill = legend_colors,
       title = "IQ Range",
       horiz = TRUE,
       cex = 0.8,
       inset = c(0, -0.16),
       xpd = TRUE) # Allow legend to be drawn outside plot area

# Add caption
mtext("Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.", 
      side = 1, line = 2.5, adj = 0.5, cex = 1, family = "serif")
```
The global map provides a broad overview of what IQ scores look like per country.


Compare Average National IQ Across All Continents
```{r Compare Average National IQ Across All Continents}
# Calculate average IQ for each continent
avg_continent_iq <- avgIQpercountry %>%
  group_by(Continent) %>%
  summarize(Average_IQ = mean(`Average IQ`, na.rm = TRUE)) %>%
  mutate(Color = sapply(Average_IQ, iq_colors), # Call global function iq_colors
         Label = sapply(Average_IQ, iq_labels)) %>%
  arrange(desc(Average_IQ)) # Sort continents in descending order of Average IQ
avg_continent_iq # print the calculated results

# Adjusting the labels for better display on the graph
avg_continent_iq$Continent <- gsub("North America", "North\nAmerica", 
                                   avg_continent_iq$Continent)
avg_continent_iq$Continent <- gsub("South America", "South\nAmerica", 
                                   avg_continent_iq$Continent)
avg_continent_iq$Continent <- gsub("Central America", "Central\nAmerica", 
                                   avg_continent_iq$Continent)

ggplot(avg_continent_iq, aes(x = reorder(Continent, -Average_IQ), y = Average_IQ, 
                             fill = Label)) +
  geom_bar(stat = "identity", color = "black") +
  scale_fill_manual(values = c("60-69" = "#FFA500", "70-79" = "#F5DEB3", 
                               "80-89" = "#ADD8E6", "90-99" = "#7B68EE"),
                    name = "IQ Level") + # Change the legend title here
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  scale_y_continuous(breaks = seq(0, 90, by = 5)) +
  labs(title = "Average National IQ Across All Continents",
       x = "Continent",
       y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title 
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1))) # Center caption
```


Bar Graph of Average IQ per Country In Africa
```{r Bar Graph of Average IQ per Country In Africa}
# Filter the dataset for African countries
africa_data <- avgIQpercountry %>% 
  filter(Continent == "Africa") %>%
  mutate(IQ_level = sapply(`Average IQ`, iq_labels))

# Create the color palette based on unique IQ levels in africa_data
iq_color_values <- sapply(unique(africa_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(africa_data$IQ_level)

# Create bar graph
ggplot(africa_data, aes(x = reorder(Country, `Average IQ`), y = `Average IQ`, 
                        fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000") +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Africa", x = "Country", 
       y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(africa_data$`Average IQ`, na.rm = TRUE), 
                                  by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(0.7)), # x-axis labels
        axis.text.y = element_text(size = rel(0.7)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title 
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In Asia
```{r Bar Graph of Average IQ per Country In Asia}
# Define the order of levels for IQ_level factor
iq_levels_ordered <- c("<50", "60-69", "70-79", "80-89", "90-99", "100+")

# Filter the dataset for Asian countries
asia_data <- avgIQpercountry %>% 
  filter(Continent == "Asia") %>%
  mutate(IQ_level = factor(sapply(`Average IQ`, iq_labels), 
                           levels = iq_levels_ordered))

# Define the color palette
iq_color_values <- c("<50" = "#8B0000", "60-69" = "#FFA500", "70-79" = "#F5DEB3", 
                     "80-89" = "#ADD8E6", "90-99" = "#7B68EE", "100+" = "#0000CD")

# Create bar graph
ggplot(asia_data, aes(x = reorder(Country, `Average IQ`), y = `Average IQ`, 
                      fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000") +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Asia", x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(asia_data$`Average IQ`, na.rm = TRUE), 
                                  by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(0.7)), # x-axis labels
        axis.text.y = element_text(size = rel(0.7)), # y-axis labels
        axis.title.x = element_text(size = rel(0.9)), # x-axis title
        axis.title.y = element_text(size = rel(0.9)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In Central America
```{r Bar Graph of Average IQ per Country In Central America}
# Filter the dataset for Central American countries
centralamerica_data <- avgIQpercountry %>% 
  filter(Continent == "Central America") %>%
  mutate(IQ_level = sapply(`Average IQ`, iq_labels))

# Create the color palette based on unique IQ levels in centralamerica_data
iq_color_values <- sapply(unique(centralamerica_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(centralamerica_data$IQ_level)

# Create bar graph
ggplot(centralamerica_data, aes(x = reorder(Country, `Average IQ`), 
                                y = `Average IQ`, fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000", width = 0.8) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Central America", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(centralamerica_data$`Average IQ`, 
                                         na.rm = TRUE), by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In Europe
```{r Bar Graph of Average IQ per Country In Europe}
# Define the order of levels for IQ_level factor
iq_levels_ordered <- c("80-89", "90-99", "100+")

# Filter the dataset for European countries
europe_data <- avgIQpercountry %>% 
  filter(Continent == "Europe") %>%
  mutate(IQ_level = factor(sapply(`Average IQ`, iq_labels), 
                           levels = iq_levels_ordered)) 

# Define the color palette
iq_color_values <- c("80-89" = "#ADD8E6", 
                     "90-99" = "#7B68EE", 
                     "100+" = "#0000CD")

# Create bar graph
ggplot(europe_data, aes(x = reorder(Country, `Average IQ`), 
                        y = `Average IQ`, fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000") +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Europe", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(europe_data$`Average IQ`, na.rm = TRUE), 
                                  by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(0.7)), # x-axis labels
        axis.text.y = element_text(size = rel(0.7)), # y-axis labels
        axis.title.x = element_text(size = rel(1)), # x-axis title
        axis.title.y = element_text(size = rel(1)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In Europe/Asia
```{r Bar Graph of Average IQ per Country In Europe/Asia}
# Filter the dataset for Eurasian countries
eurasian_data <- avgIQpercountry %>% 
  filter(Continent == "Europe/Asia") %>%
  mutate(Country = ifelse(Country == "Turkey", "Türkiye", Country), 
         IQ_level = sapply(`Average IQ`, iq_labels))

# Create the color palette based on unique IQ levels in eurasian_data
iq_color_values <- sapply(unique(eurasian_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(eurasian_data$IQ_level)

# Create bar graph
ggplot(eurasian_data, aes(x = reorder(Country, `Average IQ`), 
                          y = `Average IQ`, fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000", width = 0.5) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Europe/Asia", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(eurasian_data$`Average IQ`, na.rm = TRUE), 
                                  by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")

```


Bar Graph of Average IQ per Country In North America
```{r Bar Graph of Average IQ per Country In North America}
# Filter the dataset for North American countries
northamerican_data <- avgIQpercountry %>% 
  filter(Continent == "North America") %>%
  mutate(IQ_level = sapply(`Average IQ`, iq_labels))

# Create the color palette based on unique IQ levels in northamerican_data
iq_color_values <- sapply(unique(northamerican_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(northamerican_data$IQ_level)

# Create bar graph
ggplot(northamerican_data, aes(x = reorder(Country, `Average IQ`), 
                               y = `Average IQ`, fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000", width = 0.6) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in North America", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(northamerican_data$`Average IQ`, 
                                         na.rm = TRUE), by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In Oceania
```{r Bar Graph of Average IQ per Country In Oceania}
# Filter the dataset for Oceanian countries
oceanian_data <- avgIQpercountry %>% 
  filter(Continent == "Oceania") %>%
  mutate(IQ_level = sapply(`Average IQ`, iq_labels))  

# Create the color palette based on unique IQ levels in oceanian_data
iq_color_values <- sapply(unique(oceanian_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(oceanian_data$IQ_level)

# Create bar graph
ggplot(oceanian_data, aes(x = reorder(Country, `Average IQ`), 
                          y = `Average IQ`, fill = IQ_level)) +
  geom_bar(stat = "identity", color = "black", width = 0.8) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in Oceania", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(oceanian_data$`Average IQ`, na.rm = TRUE), 
                                  by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


Bar Graph of Average IQ per Country In South America
```{r Bar Graph of Average IQ per Country In South America}
# Filter the dataset for South American countries
southamerican_data <- avgIQpercountry %>% 
  filter(Continent == "South America") %>%
  mutate(IQ_level = sapply(`Average IQ`, iq_labels))  

# Create the color palette based on unique IQ levels in southamerican_data
iq_color_values <- sapply(unique(southamerican_data$IQ_level), iq_colors)
names(iq_color_values) <- unique(southamerican_data$IQ_level)

# Create bar graph
ggplot(oceanian_data, aes(x = reorder(Country, `Average IQ`), y = `Average IQ`, 
                          fill = IQ_level)) +
  geom_bar(stat = "identity", color = "#000000", width = 0.8) +
  coord_flip() +
  theme_minimal() +
  scale_fill_manual(values = iq_color_values, name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "Average IQ per Country in South America", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  scale_y_continuous(breaks = seq(0, max(southamerican_data$`Average IQ`, na.rm = TRUE), by = 5)) +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size = rel(1)), # x-axis labels
        axis.text.y = element_text(size = rel(1)), # y-axis labels
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 1, size = rel(1)), # Center caption
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black"), # Add black ticks on y-axis
        legend.position = "right")
```


$H_0: \mu_\text{Muslim IQ} = \mu_\text{Non-Muslim IQ}$ vs $H_a: \mu_\text{Muslim IQ} \ne \mu_\text{Non-Muslim IQ}$


Hypothesis Testing 2
```{r Hypothesis Testing 2}
muslim_nonmuslim <- avgIQpercountry %>% 
  filter(Country %in% c("United States", "Israel", "Palestine", "Syria", 
                        "Jordan", "Egypt", "Lebanon", "Iraq"))
muslim_nonmuslim$Group <- ifelse(muslim_nonmuslim$Country %in% 
                                     c("United States", "Israel"), "Non-Muslim", 
                                                                   "Muslim")
# Perform a One-Way Analysis of Variance (ANOVA)
abab_israeli <- aov(`Average IQ`~Group, data = muslim_nonmuslim)
summary(abab_israeli)
TukeyHSD(abab_israeli) # pairwise comparison test
```
f = 13.16; p-value = 0.011

Reject H0; there is no evidence to support the claim that the mean IQs between
these Muslim and Non-Muslim countries are equal at the 0.05 significance level.

In the pairwise comparison test, it is found the Muslim and Non-Muslim countries 
are significantly different in average IQ scores. This is denoted by the low 
probability value (i.e. 0.010999) in the "p adj" column. Viz., the probability 
this pair is equivalent is 1.0999%.


Compare IQ in Countries Involved in Arab-Israeli Conflict
```{r Compare IQ in Countries Involved in Arab-Israeli Conflict}
# Define the grouping of countries and prepare the data
selected_countries_data <- avgIQpercountry %>%
  filter(Country %in% c("United States", "Israel", 
                        "Palestine", "Syria",
                        "Jordan", "Egypt",
                        "Lebanon", "Iraq")) %>%
  select(Country, `Average IQ`) %>%
  arrange(desc(`Average IQ`)) %>% # Sorting by descending IQ
  mutate(Color = sapply(`Average IQ`, iq_colors), # Apply the iq_colors function
         IQ_Range = sapply(`Average IQ`, iq_labels), # Apply the iq_labels function
         Category = factor(ifelse(Country %in% c("United States", "Israel"), 
                                  "Non-Muslim", "Muslim"), 
                           levels = c("Non-Muslim", "Muslim")))

# Ensure the order of factors in Country matches the descending IQ order
selected_countries_data$Country <- factor(selected_countries_data$Country, 
                                          levels = selected_countries_data$Country)

# Graph their IQs with custom colors and relative text sizing
ggplot(selected_countries_data, aes(x = Country, y = `Average IQ`, 
                                    fill = IQ_Range)) +
  geom_bar(stat = "identity", color = "#000000") +
  scale_fill_manual(values = setNames(as.character(selected_countries_data$Color), 
                                      selected_countries_data$IQ_Range), 
                    name = "IQ Level") +
  guides(fill = guide_legend(reverse = TRUE)) + # Reverse legend order
  labs(title = "IQ Scores of Participants in Arab-Israeli Conflict", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(hjust = 0.5, size = rel(0.8)), # x-axis label
        axis.text.y = element_text(size = rel(1.1)), # y-axis label
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 0.5, size = rel(0.8)),# Center caption
        strip.text = element_text(size = rel(1.2)), # facet titles
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black")) + # Add black ticks on y-axis
  scale_y_continuous(breaks = seq(0, 100, by = 10)) + # y-axis ticks
  facet_wrap(~ Category, scales = "free_x") # Split graph into two categories
```


$H_0: \mu_\text{Allied With Russia IQ} = \mu_\text{NATO IQ}$ vs $H_a: \mu_\text{Allied With Russia IQ} \ne \mu_\text{NATO IQ IQ}$


Hypothesis Testing 3
```{r Hypothesis Testing 3}
# Create the dataset and assign groups
NATO_nonNATO <- avgIQpercountry %>%
  filter(Country %in% c("Russia", "Syria", "Belarus", "Ukraine", "United States", 
                        "United Kingdom", "Germany", "France", "Poland", 
                        "Canada", "Turkey", "Estonia", "Latvia", "Lithuania",
                        "Moldova", "Hungary", "Slovakia")) %>%
  mutate(Group = ifelse(Country %in% c("Russia", "Belarus", "Syria"), 
                        "Allied With Russia", 
                        "NATO"))

# Perform a One-Way Analysis of Variance (ANOVA)
abab_israeli2 <- aov(`Average IQ`~Group, data = NATO_nonNATO)
summary(abab_israeli2)
TukeyHSD(abab_israeli2) # pairwise comparison test
```
f = 1.786; p-value = 0.203

Fail to reject H0; there is no evidence to support the claim that the mean IQs 
between these NATO and Non-NATO countries are different at the 0.05 significance 
level.

In the pairwise comparison test, it is found that NATO and Non-NATO countries 
are not significantly different in average IQ scores. This is denoted by the 
high probability value (i.e. 0.2027295) in the "p adj" column. Viz., the 
probability this pair is equivalent is 20.27%.


Compare IQ in Countries Involved in Russo-Ukrainian War
```{r Compare IQ in Countries Involved in Russo-Ukrainian War}
# Define the grouping of countries and prepare the data
selected_countries_data <- avgIQpercountry %>%
  filter(Country %in% c("Russia", "Syria", "Belarus", "Ukraine", "United States", 
                        "United Kingdom", "Germany", "France", "Poland", 
                        "Canada", "Turkey", "Estonia", "Latvia", "Lithuania",
                        "Moldova", "Hungary", "Slovakia")) %>%
  mutate(Country = gsub("Turkey", "Türkiye", Country)) %>% # Replace Turkey with Türkiye
  select(Country, `Average IQ`) %>%
  arrange(desc(`Average IQ`)) %>% # Sorting by descending IQ
  mutate(Color = sapply(`Average IQ`, iq_colors), # Apply the iq_colors function
         IQ_Range = sapply(`Average IQ`, iq_labels), # Apply the iq_labels function
         Category = factor(ifelse(Country %in% c("Russia", "Belarus", "Syria"), 
                                  "Allied with Russia", "European Union/NATO"),
                           levels = c("Allied with Russia", "European Union/NATO")))

# Ensure the order of factors in Country matches the descending IQ order
selected_countries_data$Country <- factor(selected_countries_data$Country, 
                                          levels = selected_countries_data$Country)

# Graph their IQs with custom colors and relative text sizing
ggplot(selected_countries_data, aes(x = Country, y = `Average IQ`, 
                                    fill = IQ_Range)) +
  geom_bar(stat = "identity", color = "#000000") +
  scale_fill_manual(values = setNames(as.character(selected_countries_data$Color), 
                                      selected_countries_data$IQ_Range),
                    name = "IQ Level", 
                    breaks = c("100+", "90-99", "80-89", "70-79")) + # Order legend
  guides(fill = guide_legend(reverse = FALSE)) + # Maintain specified order
  labs(title = "IQ Scores of Countries in the Russo-Ukrainian War", 
       x = "Country", y = "Average IQ",
       caption = "Source: Kaggle/ Google LLC (2023) Average Global IQ per Country with Other Stats.") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(angle = 90, hjust = 1, 
                                   vjust = .5, size = rel(0.8)), # x-axis label
        axis.text.y = element_text(size = rel(1.1)), # y-axis label
        axis.title.x = element_text(size = rel(1.2)), # x-axis title
        axis.title.y = element_text(size = rel(1.2)), # y-axis title
        plot.title = element_text(hjust = 0.5, size = rel(1.5)), # Center title
        plot.caption = element_text(hjust = 0.5, size = rel(0.8)), # Center caption
        strip.text = element_text(size = rel(1.2)), # facet titles
        axis.ticks.x = element_line(color = "black"), # Add black ticks on x-axis
        axis.ticks.y = element_line(color = "black")) + # Add black ticks on y-axis
  scale_y_continuous(breaks = seq(0, 100, by = 10)) + # y-axis ticks
  facet_wrap(~ Category, scales = "free_x") # Split graph into two categories
```
